Advanced Normalization Tools (ANTS)

### 2010-02-05 Documentation/Tex/ants.tex
aff description
git-svn-id: https://advants.svn.sourceforge.net/svnroot/advants@307 fd2da0a4-d968-48df-9cb5-d8c496287cbf
ants aff explanation
git-svn-id: https://advants.svn.sourceforge.net/svnroot/advants@308 fd2da0a4-d968-48df-9cb5-d8c496287cbf

\subsection{I/O Data Formats In ANTS}
ANTS support both 2D and 3D images. 
Since ANTS is implemented in ITK, 
it is able to read and write the popular data formats supported through ITK. 
There are four basic types of data for ANTS.
\subsubsection{Image Volumes}
ANTS support both 2D and 3D images, including
\begin{itemize}
\item Nifti (.nii, .nii.gz)
\item Analyze (.hdr + .img / .img.gz)
\item MetaImage (.mha)
\item Other formats through itk::ImageFileWriter / itk::ImageFileWriter such as jpg, tiff, etc.  See ITK documentation. 
\end{itemize}
\subsubsection{Affine Transformation File}
ANTS uses the itk::TransformFileReader / itk::TransformFileWriter to handle affine transformation files. 
These are text files recording the matrix and offset for the affine transform {\bf FIXME NEED MORE DETAIL on definition and geometry of affine matrix}.
\subsubsection{Deformation Field File}
ANTS writes the deformation field using one file for the deformation in each dimension. 
For example, if the deformation field is named ABCD.nii.gz and exists in three dimensions, the corresponding files are:
ABCDxvec.nii.gz, 
ABCDyvec.nii.gz, 
ABCDzvec.nii.gz (for 3D).  Note that ANTS programs {\em assume} the $*$vec part of the name will exist.  Thus, 
one refers to the warp file as  ABCD.nii.gz , dropping the $*$vec from the name. 
\subsubsection{Labeled Point Sets}
There are three formats supported
\begin{itemize}
\item  VTK format: point + scalar label in a .vtk file. {\bf FIXME NEED MORE DETAIL -- what type of vtk data ?  polydata?}.
\item  Labeled point sets are saved as an image volume with the intensity = 1,2,3 for label = 1,2,3. For example, 
figure~\ref{fig:frown} shows 2D images representing a 3-class labeled point set. 
\item Avants Format. In this simple format, the point sets are saved as a text file. The first and last line is always 0 0 0 0. Each line in between for a point as X Y Z L (L is the label.)
\end{itemize}
The names and formats of the output files are both specified by -o option to the ANTS executable. 
For example, for in 3D with -o RegA2B.nii.gz , the typical output files are:
Affine transform file: 
RegA2BAffine.txt 
Warping deformation field: 
RegA2BWarpxvec.nii.gz, RegA2BWarpyvec.nii.gz, RegA2BWarpzvec.nii.gz
Inverse warping deformation field:
RegA2BInverseWarpxvec.nii.gz, RegA2BInverseWarpyvec.nii.gz, RegA2BInverseWarpzvec.nii.gz.  
To apply the affine transform and the deformation field to the moving image, use WarpImageMultiTransform. 
For example: to apply the transform to the moving image:

>>>>

\subsection{I/O Data Formats In ANTS}
ANTS support both 2D and 3D images. 
Since ANTS is implemented in ITK, 
it is able to read and write the popular data formats supported through ITK. 
There are four basic types of data for ANTS.
\subsubsection{Image Volumes}
ANTS support both 2D and 3D images, including
\begin{itemize}
\item Nifti (.nii, .nii.gz)
\item Analyze (.hdr + .img / .img.gz)
\item MetaImage (.mha)
\item Other formats through itk::ImageFileWriter / itk::ImageFileWriter such as jpg, tiff, etc.  See ITK documentation. 
\end{itemize}
\subsubsection{Affine Transformation File}
ANTS uses the itk::TransformFileReader / itk::TransformFileWriter to handle affine transformation files. 
These are text files recording the matrix and offset for the affine transform {\bf FIXME NEED MORE DETAIL on definition and geometry of affine matrix}.
The interpretation depends on the header but generally it breaks down like this:

\#Insight Transform File V1.0

\# Transform 0

Transform: MatrixOffsetTransformBase\_double\_3\_3

Parameters:  $ a~b~c~d~e~f~g~h~i~j~k~l $

FixedParameters: $ M~N~O $

The meaning of these different terms:

The matrix component (rotation, scaling, shearing all composed together):

$ \begin{matrix}
a & b & c \\
d & e & f  \\
g & h & i \\ 
\end{matrix} $

is followed by  $~j~k~l~$ --- this is the translation.

$L~M~N$  is the center of rotation .

You can use the --initial-affine  and    --continue-affine  options to control how this is used in registration.

\subsubsection{Deformation Field File}
ANTS writes the deformation field using one file for the deformation in each dimension. 
For example, if the deformation field is named ABCD.nii.gz and exists in three dimensions, the corresponding files are:
ABCDxvec.nii.gz, 
ABCDyvec.nii.gz, 
ABCDzvec.nii.gz (for 3D).  Note that ANTS programs {\em assume} the $*$vec part of the name will exist.  Thus, 
one refers to the warp file as  ABCD.nii.gz , dropping the $*$vec from the name. 
\subsubsection{Labeled Point Sets}
There are three formats supported
\begin{itemize}
\item  VTK format: point + scalar label in a .vtk file. {\bf FIXME NEED MORE DETAIL -- what type of vtk data ?  polydata?}.
\item  Labeled point sets are saved as an image volume with the intensity = 1,2,3 for label = 1,2,3. For example, 
figure~\ref{fig:frown} shows 2D images representing a 3-class labeled point set. 
\item Avants Format. In this simple format, the point sets are saved as a text file. The first and last line is always 0 0 0 0. Each line in between for a point as X Y Z L (L is the label.)
\end{itemize}
The names and formats of the output files are both specified by -o option to the ANTS executable. 
For example, for in 3D with -o RegA2B.nii.gz , the typical output files are:
Affine transform file: 
RegA2BAffine.txt 
Warping deformation field: 
RegA2BWarpxvec.nii.gz, RegA2BWarpyvec.nii.gz, RegA2BWarpzvec.nii.gz
Inverse warping deformation field:
RegA2BInverseWarpxvec.nii.gz, RegA2BInverseWarpyvec.nii.gz, RegA2BInverseWarpzvec.nii.gz.  
To apply the affine transform and the deformation field to the moving image, use WarpImageMultiTransform. 
For example: to apply the transform to the moving image:
